---
title: "Matching drug_profiles_filtered_Samples based on Mahalanobis Distance Metric"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(tidyverse)
library(knitr)
library(magrittr)
library(ggplot2)
library(corrplot)
library(cytominer)
library(data.table)
library(RColorBrewer)

```

# Importing Combined_drug_target Dataset

```{r}
# Path of combined data

path <- "/Users/habbasi/Desktop/2016_03_14_TargetID_Wagner_Schenone_BWH/similarity_matrix/combined_median_target_Drug_CSV/combined_normalized_variable_selected/normalized_variable_selected_drug_target_MD_95th_percentile.csv"

out <- "/Users/habbasi/Desktop/2016_03_14_TargetID_Wagner_Schenone_BWH/similarity_matrix/outputs/012220_matcing_drugs_profiles_MahlanobisDistance/95th_percentile"

moa_path <- "/Users/habbasi/Desktop/2016_03_14_TargetID_Wagner_Schenone_BWH/metadata/target_ID_MoA.csv"
```

# Importing combined data
```{r}

data <- data.frame(readr::read_csv(path), stringsAsFactors = F)
data$X1 <- NULL


# reading and renaming column names and selecting substring of compound names and three Metadatacolumns

moa <- readr::read_csv(moa_path) %>% 
   setnames(old= c("BRD", "submitter"), new=c("Metadata_broad_sample","Metadata_submitter")) %>%    
   select(Metadata_broad_sample, Metadata_submitter) %>% 
   mutate(Metadata_platemap = "t") %>%
   mutate(Metadata_broad_sample = str_sub(Metadata_broad_sample, 1, 13))
```



# Defining Variables 

```{r}
# Metadata variables
metadata <- colnames(data) %>%
  stringr::str_subset(("^Metadata_"), negate = FALSE)



# Variables
variables <- colnames(data) %>%
 str_subset(("^Metadata"), negate=TRUE)
```


# Making a new column Metadata_platemap and mutating column values if treatments are from drug or from target data
```{r}


final <- data %>%
   mutate(Metadata_platemap= ifelse((Metadata_Plate_Map_Name == "C-7217-01-B80-003-A549"),"t", "d")) %>%
   mutate(Metadata_platemap= ifelse((Metadata_broad_sample == "DMSO"),"DMSO", Metadata_platemap)) 


# Defining new Variables
variables <- colnames(final) %>%
 str_subset(("^Metadata"), negate=TRUE) 
metadata <- colnames(final) %>%
   str_subset(("^Metadata"), negate=FALSE)


```

# Filtering out Samples based on 95th percentile Mahalanobis distance Assignment

```{r}

final_filtered <- final %>%
   filter(Metadata_Assignment == TRUE)

```


# Median Aggregation of compound Profiles

```{r}


# Mean Aggregation of compound profiles for each concentration

aggregated <- cytominer::aggregate(population = final_filtered,
                        variables = variables,
                        strata =  c("Metadata_broad_sample", "Metadata_platemap",   
                                    "Metadata_mmoles_per_liter"),
                        operation = "median"
                        )
aggregated$X1 <- NULL


# Joining moa list with metadata columns

pmeta <- final %>% 
  select(metadata) %>%
  left_join(moa, by= c("Metadata_broad_sample", "Metadata_platemap")) %>%
  select(Metadata_broad_sample, Metadata_moa,Metadata_pert_name, Metadata_platemap, Metadata_submitter) %>%
   distinct(Metadata_broad_sample, Metadata_platemap, .keep_all = TRUE) 



```

# Joining of metadata columns with mean aggregated compound profiles
```{r}


prf <- aggregated %>% 
   left_join(pmeta, by= c("Metadata_broad_sample", "Metadata_platemap"))


# metadata variables

metadata <- colnames(prf) %>%
   str_subset(("^Metadata"), negate=FALSE) 
  

# Variables
variables <- colnames(prf) %>%
 str_subset(("^Metadata"), negate=TRUE)

```

# similarity matrix

```{r}
# get data matrix
 data_matrix <-
   prf %>%
   select(variables)



 metadata <-
   prf %>%
     select(matches("Metadata")) %>% 
     mutate(Metadata_broad_sample = str_sub(Metadata_broad_sample, 1, 13)) %>%
    unite(Metadata_sample, c("Metadata_broad_sample", "Metadata_platemap", "Metadata_mmoles_per_liter")) %>%
    mutate(Metadata_sample = str_replace(Metadata_sample, "DMSO_DMSO", "DMSO"), Metadata_sample =       
              str_replace(Metadata_sample, "BRD-", " ")) %>%
    rowid_to_column(var = "id")


```


# Calculating Similarity Profiles

```{r}
similarity <- cor(t(data_matrix))

row.names(similarity) <- metadata$Metadata_sample
colnames(similarity) <- metadata$Metadata_sample



#write.csv(similarity, paste0(out, "/", "similarity_correlations_MD.csv"))


# Selecting upper half of correlations and setting the diagonal and lower halt to NA 
upper_tri <- function(cr) {
  cr[lower.tri(cr)] = NA
  cr[which(cr == 1)] = NA
  return (cr)
}

cr.melt <- upper_tri(similarity)

row.names(cr.melt) <- metadata$Metadata_sample
colnames(cr.melt) <- metadata$Metadata_sample

# Rehaping of matrix and removing lower and diagnal NA's

cr.final <- cr.melt  %>% 
   reshape2::melt() %>% 
   na.omit() %>%
   inner_join(
     metadata,
     by = c("Var1" = "Metadata_sample")
   ) %>%
   rename(
     Metadata_Compound1 = Var1,
     Metadata_MoA1 = Metadata_moa,
     Metadata_pert_name1 = Metadata_pert_name,
     Metadata_submitter1 = Metadata_submitter,
     correlation = value) %>% 
   inner_join(metadata,
     by = c("Var2" = "Metadata_sample")
   ) %>% 
   rename(
     Metadata_Compound2 = Var2,
     Metadata_MoA2 = Metadata_moa,
     Metadata_pert_name2 = Metadata_pert_name,
     Metadata_submitter2 = Metadata_submitter) %>% arrange(desc(correlation))

# Removing index column

cr.final[, c('id.x', 'id.y')] <- NULL


# Calculating quantile distribution of correlations

quantile_dist <- quantile(cr.final$correlation, prob = c(0.05, 0.95))


bottom5_threshold <- quantile_dist[1]
top5_threshold <- quantile_dist[2]

top5_percent <- cr.final[which(cr.final$correlation >= top5_threshold), ]
bottom5_percent <- cr.final[which(cr.final$correlation <= bottom5_threshold), ]


#write.csv(top5_percent, paste0(out, "/", "top5percent_correlations_MD.csv"))
#write.csv(bottom5_percent,  paste0(out, "/", "bottom5percent_correlations_MD.csv"))

# Excluding self drugrepurposing and target_ID correlations

indx1 <- with(top5_percent, grepl("_d", Metadata_Compound1))
indx2 <- with(top5_percent, grepl("_d", Metadata_Compound2))
drug_top5combinations <- top5_percent[indx1&indx2, ]

indx3 <- with(top5_percent, grepl("_t", Metadata_Compound1))
indx4 <- with(top5_percent, grepl("_t", Metadata_Compound2))
target_top5combinations <- top5_percent[indx3&indx4, ]


top_united <- union(drug_top5combinations, target_top5combinations)
target_top5percent <- setdiff(top5_percent, top_united)
```

# Restricting Target_ID drugs to first column
```{r}
   
target_filtered <- target_top5percent[grep("_t", target_top5percent$Metadata_Compound1), ]


# Filtering out drugsRepurposing drugs in first column

oldnames = c("Metadata_Compound1","Metadata_Compound2", "correlation", "Metadata_MoA1","Metadata_pert_name1", "Metadata_submitter1", "Metadata_MoA2",       "Metadata_pert_name2", "Metadata_submitter2")
newnames = c("Metadata_Compound2","Metadata_Compound1", "correlation", "Metadata_MoA2","Metadata_pert_name2", "Metadata_submitter2", "Metadata_MoA1",       "Metadata_pert_name1", "Metadata_submitter1")
drug_filtered <- target_top5percent[grep("_d", target_top5percent$Metadata_Compound1, perl=TRUE), ] %>% 
   rename_at(vars(oldnames), ~newnames)
drug_filtered <- drug_filtered[, oldnames]

top5percent_profiles <- rbind(target_filtered, drug_filtered)
top5percent_profiles <- top5percent_profiles[order(top5percent_profiles$Metadata_pert_name1, decreasing = TRUE), ]

write.csv(top5percent_profiles, paste0(out, "/", "target_top5percent_hits_MD.csv"))

```


```{r}


# Excluding self drugrepurposing and target_ID correlations

indx5 <- with(bottom5_percent, grepl("_d", Metadata_Compound1))
indx6 <- with(bottom5_percent, grepl("_d", Metadata_Compound2))
drug_bottom5combinations <- bottom5_percent[indx5&indx6, ]

indx7 <- with(bottom5_percent, grepl("_t", Metadata_Compound1))
indx8 <- with(bottom5_percent, grepl("_t", Metadata_Compound2))
target_bottom5combinations <- bottom5_percent[indx7&indx8, ]

bottom_united <- union(drug_bottom5combinations,target_bottom5combinations)
target_bottom5percent <- setdiff(bottom5_percent, bottom_united)



target_bottom_filtered <- target_bottom5percent[grep("_t", target_bottom5percent$Metadata_Compound1), ]


# Filtering out drugsRepurposing drugs in first column

oldnames = c("Metadata_Compound1","Metadata_Compound2", "correlation", "Metadata_MoA1","Metadata_pert_name1", "Metadata_submitter1", "Metadata_MoA2",       "Metadata_pert_name2", "Metadata_submitter2")
newnames = c("Metadata_Compound2","Metadata_Compound1", "correlation", "Metadata_MoA2","Metadata_pert_name2", "Metadata_submitter2", "Metadata_MoA1",       "Metadata_pert_name1", "Metadata_submitter1")

drug_bottom_filtered <- target_bottom5percent[grep("_d", target_bottom5percent$Metadata_Compound1, perl=TRUE), ] %>% 
   rename_at(vars(oldnames), ~newnames)

# Ordering of second dataframe columns similar to the first one

drug_bottom_filtered <- drug_bottom_filtered[, oldnames]


# Merging of both dataframes together and sorting row according to Metadata_pert_name1 column

bottom5percent_profiles <- rbind(target_bottom_filtered, drug_bottom_filtered)
bottom5percent_profiles <- bottom5percent_profiles[order(bottom5percent_profiles$Metadata_pert_name1, decreasing = TRUE), ]

write.csv(bottom5percent_profiles, paste0(out, "/", "target_bottom5percent_hits_MD.csv"))

```


