---
title: "Mahalanobis Distance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(knitr)
library(magrittr)
library(ggplot2)
library(cytominer)
library(caret)
library(matrixStats)
library(wesanderson)
```

# Importing Combined_drug_target Dataset

```{r}
# Path of combined data

path<- "/Users/habbasi/Desktop/2016_03_14_TargetID_Wagner_Schenone_BWH/similarity_matrix/combined_median_target_Drug_CSV/combined_normalized_variable_selected/normalized_variable_selected_drug_target.csv"


out <- "/Users/habbasi/Desktop/2016_03_14_TargetID_Wagner_Schenone_BWH/similarity_matrix/combined_median_target_Drug_CSV/combined_normalized_variable_selected"



figpath <- "/Users/habbasi/Desktop/2016_03_14_TargetID_Wagner_Schenone_BWH/Github_repo/2016_03_14_TargetID_Wagner_Schenone_new/Figures/012020/"
```


# Mahalanobis Distance Calculated using Normalized features
# Importing normalized data

```{r}

data_normalized <- data.frame(readr::read_csv(path), stringsAsFactors = F)
data_normalized$X1 <- NULL

```

# Defining Variables 

```{r}
# Metadata variables
metadata <- colnames(data_normalized) %>%
  stringr::str_subset(("^Metadata_"), negate = FALSE)


# Variables
variables <- colnames(data_normalized) %>%
 str_subset(("^Metadata"), negate=TRUE)

```

# Filtering out trouble making variables as they have determinant zero and inverse of covariance matrix cannot be calculated
```{r}
var <- c("Cells_Neighbors_SecondClosestDistance_Adjacent", "Cytoplasm_AreaShape_MaxFeretDiameter", "Cytoplasm_AreaShape_MinFeretDiameter","Cytoplasm_Intensity_MinIntensityEdge_DNA", "Cytoplasm_Intensity_MinIntensity_DNA", "Cells_Neighbors_AngleBetweenNeighbors_Adjacent")

df <- data_normalized %>% 
  select(-var)

# Filtering out DMSO cells for both Drug Repurposing and Target datasets for calculating Mahlanobis distance on DMSO's

df_dmso <- df %>% 
  filter(Metadata_broad_sample_type == "control")


# Defining new Variables
variables <- colnames(df) %>%
 str_subset(("^Metadata"), negate=TRUE) 

# Calculating Mahlanobis distance of controls separately and taking the 85th percentile of MD values are taken as a cutoff

df_dmso$Metadata_MD <- mahalanobis(df_dmso[, variables], colMeans(df_dmso[, variables]), cov(df_dmso[, variables]))
df_dmso$Metadata_MD %>% round(2)

control_threshold <- df_dmso %>% 
   select(Metadata_MD) %>% 
   as.matrix() %>% 
  quantile(., c(.85)) %>%
  round(2)


label <- c("Total no of Samples: 3614 \n \n Threshold value \n 85th Percentile value: 1129.38")
g1 <- ggplot(data=df_dmso, aes(df_dmso$Metadata_MD)) + 
  geom_histogram(
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  geom_vline(aes(xintercept=control_threshold), 
                                          color="black", 
                                          linetype="dashed") +
   geom_text(aes(x=1800, y=700),
           label= label,hjust=0, size=4, color="#636363") +
  annotate("rect",xmin=0,
                  xmax = control_threshold,ymin=0,
                  ymax=900, alpha=.2, fill="red") +
  labs(x="Distances", y="Frequency") +
     ggtitle(label = "Distribution of Mahalanobis Distances of DMSO's") + theme(
  plot.title = element_text(hjust = 0.5)
)
 

g1
ggsave(g1, filename=paste0(figpath, "/", "MD_Z_normalized_features_DMSO_85th_percentile.png"))
```


# Calculating Mahlanobis Distance of samples
```{r}

df$Metadata_MD <- mahalanobis(df[, variables], colMeans(df[, variables]), cov(df[, variables]))
df$Metadata_MD %>% round(2)


df_trt <- df %>% filter(Metadata_broad_sample_type == "trt")

# Adding text to plot

label <- c("14.144% (7658 out of 54143) samples retained \n \n Treatments retained (85 percentile thresholding) \n DrugRepurposing: 528 out of 1570 \n Target_ID: 36 out of 81")

 g2 <- ggplot(data=df_trt, aes(df_trt$Metadata_MD)) + 
  geom_histogram(
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
    geom_text(aes(x=15000, y=35000),
           label= label,hjust=0, size=4, color="#636363") +
  geom_vline(aes(xintercept=control_threshold), 
                                          color="black", 
                                          linetype="dashed") +
  annotate("rect",xmin=-1,
                  xmax = control_threshold,ymin=0,
                  ymax=41000, alpha=.2, fill="blue") +
  labs(x="Distances", y="Frequency") +
     ggtitle(label = "Distribution of Mahalanobis Distances of Treatments") + theme(
  plot.title = element_text(hjust = 0.5)
)
g2

ggsave(g2, filename=paste0(figpath, "/", "MD_Z_normalized_features_Treatments_85th_percentile.png"))

```


# Keeping only the Treatments which are above the threshold which is 85th percentile of DMSO's Mahalanobis Distance

```{r}

df$Metadata_Assignment <- FALSE
df$Metadata_Assignment[df$Metadata_MD > control_threshold] <- TRUE


write.csv(df, paste0(out, "/", "normalized_variable_selected_drug_target_MD_85th_percentile.csv"))



```

